---
title: "Diner"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(readr)
search()
```


```{r}
members <- read.csv("members.csv",fileEncoding="UTF-8-BOM")
members
```

```{r}
members <- members %>%
  mutate(join_date=as.Date(join_date, format = "%d-%m-%Y"))
members
```

```{r}
menu <- read.csv("menu.csv",fileEncoding="UTF-8-BOM")
menu
```


```{r}
sales <- read.csv("sales.csv",fileEncoding="UTF-8-BOM")
sales
```
```{r}
sales <- sales %>%
  mutate(order_date=as.Date(order_date, format = "%d-%m-%Y"))
sales
```


```{r}
sales_menu <- merge(sales,menu,by.x = "product_id",by.y = "product_id")
head(sales_menu)
```

```{r}
all <- merge(sales_menu,members,by.x = "customer_id",by.y = "customer_id")
all
```


Question 1. What is the total amount each customer spent at the restaurant?

```{r}
library(dplyr)
amount_spent <- sales_menu %>%
  group_by(customer_id) %>%
  summarise(total_amount = sum(price)) %>%
  ungroup() %>%
  arrange(desc(total_amount))
amount_spent
```
Question 2. How many days has each customer visited the restaurant?

```{r}
no_of_visits <- sales %>%
  group_by(customer_id) %>%
  summarise(no = length(unique(order_date))) %>%
  ungroup()
no_of_visits
```
Question 3. What was the first item from the menu purchased by each customer?

```{r}
first_item <- sales_menu %>%
  arrange(customer_id,order_date) %>%
  group_by(customer_id) %>%
  summarise(first_item = first(product_name))
first_item

first_item <- sales_menu %>%
  arrange(customer_id,order_date) %>%
  group_by(customer_id) %>%
  filter(order_date == min(order_date)) %>%
  ungroup() %>%
  select(customer_id,product_name) 
unique(first_item[,c("customer_id","product_name")])

```
Question 4. What is the most purchased item on the menu and how many times was it purchased by all customers?

```{r}
most_purchased <- sales_menu %>%
  group_by(product_name) %>%
  summarise(number = n()) %>%
  ungroup() %>%
  filter(number == max(number))
most_purchased
```

Question 5. Which item was the most popular for each customer?

```{r}
most_pop <- sales_menu %>%
  group_by(customer_id,product_name) %>%
  summarise(ct = n()) %>%
  filter(ct == max(ct)) %>%
  ungroup() 
most_pop
```

Question 6. Which item was purchased first by the customer after they became a member?

```{r}

first_purchase <- all %>%
  filter(order_date >= join_date) %>%
  arrange(customer_id,order_date) %>%
  group_by(customer_id) %>%
  filter(order_date == min(order_date)) %>% #incase more than 1 item was purchased in same order
  ungroup() %>%
  select(customer_id,order_date,product_name,join_date)
first_purchase
  
```

Question 7. Which item was purchased just before the customer became a member?

```{r}
last_purchase <- all %>%
  filter(order_date < join_date) %>%
  group_by(customer_id) %>%
  filter(order_date == max(order_date)) %>% #incase more than 1 item was purchased in same order
  ungroup() %>%
  select(customer_id,order_date,product_name,join_date)
last_purchase
```
Question 8. What is the total items and amount spent for each member before they became a member?

```{r}
items_spent <- all %>%
  filter(order_date < join_date) %>%
  group_by(customer_id) %>%
  summarise(total_spend = sum(price),number_of_item = n()) %>%
  ungroup() 
items_spent
```

Question 9. If each $1 spent equates to 10 points and sushi has a 2x points multiplier - how many points would each customer have

```{r}
points <- sales_menu %>%
  mutate(point = if_else(product_name == "sushi", 20*price,10*price )) %>%
  group_by(customer_id) %>%
  summarise(total_points = sum(point)) %>%
  ungroup() %>%
  arrange(desc(total_points))
points
```

Question 10. In the first week after a customer joins the program (including their join date) they earn 2x points on all items, not just sushi - how many points do customer A and B have at the end of January?

```{r}
new_points <- all %>%
  filter(order_date < "2021-01-31") %>%
  mutate(point = case_when((order_date >= join_date) & (order_date <join_date+7) ~ (20*price),
                           product_name == "sushi"~ 20*price,
                           TRUE ~ 10*price)) %>%
  group_by(customer_id) %>%
  summarise(total_points = sum(point)) %>%
  ungroup() %>%
  arrange(desc(total_points))
new_points
```


Question 11. Recreate the following table
customer_id	order_date	product_name	price	member
A	2021-01-01	curry	15	N
A	2021-01-01	sushi	10	N
A	2021-01-07	curry	15	Y
A	2021-01-10	ramen	12	Y
A	2021-01-11	ramen	12	Y
A	2021-01-11	ramen	12	Y
B	2021-01-01	curry	15	N
B	2021-01-02	curry	15	N
B	2021-01-04	sushi	10	N
B	2021-01-11	sushi	10	Y
B	2021-01-16	ramen	12	Y
B	2021-02-01	ramen	12	Y
C	2021-01-01	ramen	12	N
C	2021-01-01	ramen	12	N
C	2021-01-07	ramen	12	N

```{r}
new <- left_join(sales,members,by="customer_id")
new <- merge(new,menu,by.x = "product_id",by.y = "product_id")
new <- new %>%
  mutate(member = case_when((order_date >= join_date) ~ "Y",
                            order_date<join_date ~ "N",
                            TRUE ~ "N")) %>%
  select(customer_id,order_date,product_name,price,member) %>%
  arrange(customer_id,order_date,product_name)
new
```

Question 12) Danny also requires further information about the ranking of customer products, but he purposely does not need the ranking for non-member purchases so he expects null ranking values for the records when customers are not yet part of the loyalty program.

```{r}
rank1 <- new %>%
  filter(member=="Y") %>%
  group_by(customer_id) %>%
  mutate(ranks = rank(order_date, ties.method = "min"))
rank2 <- new %>%
  filter(member =="N") %>%
  mutate(ranks = NULL)

rank = rbind(rank1, rank2)
rank <= rank %>%
  arrange(customer_id)
rank
```



